---
title: "Exploratory analysis: gender and cognitive competence"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
warning: false
editor: visual
embed-resources: true
---

### Attatching packages
```{r}
#| message: false

library(readr)
library(tidyverse)
```

### Loading in data
```{r}
#| message: false
allYrsFinal <- read_csv("FinalFiles2023/Data/All Years Final Public - with Vars.csv")
```

### Gender and Pre iscamboxplot
```{r}
iscamboxplot(allYrsFinal$cognitive.competence.pre, explanatory=allYrsFinal$gender)
```

### Gender and Pre violin + boxplot

```{r}
#cognitive.competence.pre	cognitive.competence.post
#cognitive.competence.change	
# gender
allYrsFinal|>
  #filter(!is.na(gender), !is.na(cognitive.competence.pre)) |>
  ggplot(aes(x=gender, y=cognitive.competence.pre)) + 
  geom_boxplot() + geom_violin(alpha = 0.5, color = "orchid")
```

### Gender and Pre t-test
```{r}
t.test(allYrsFinal$cognitive.competence.pre ~ allYrsFinal$gender)
```
### Gender and Post iscamboxplot
```{r}
iscamboxplot(allYrsFinal$cognitive.competence.post, explanatory=allYrsFinal$gender)
```

### Gender and Post violin + boxplot
```{r}
allYrsFinal|>
  #filter(!is.na(gender), !is.na(cognitive.competence.post)) |>
  ggplot(aes(x=gender, y=cognitive.competence.post)) + 
  geom_boxplot() + geom_violin(alpha = 0.5, color="orchid")
```

### Gender and Post t-test
```{r}
t.test(allYrsFinal$cognitive.competence.post ~ allYrsFinal$gender)
```

### Gender and Change iscamboxplot
```{r}
iscamboxplot(allYrsFinal$cognitive.competence.change, explanatory=allYrsFinal$gender)
```


### Gender and Change boxplot
```{r}
allYrsFinal|>
  #filter(!is.na(gender), !is.na(cognitive.competence.change)) |>
  ggplot(aes(x=gender, y=cognitive.competence.change)) + 
  geom_boxplot() + geom_violin(alpha = 0.5, color = "orchid")
```

### Gender and Change t-test
```{r}
t.test(allYrsFinal$cognitive.competence.change ~ allYrsFinal$gender)
```

## By Question t-tests
Post:

-   6e: I had trouble understanding statistics because of how I think.

-   7a: I had no idea of what was going on in this statistics course.

-   8f: I made a lot of math errors in statistics.

-   9a: I learned statistics.

-   9b: I understood statistics equations.

-   9e: I found it difficult to understand statistical concepts.

t-tests per question:
```{r}
competence <- c("^q6e", "^q7a", "^q8f", "^q9a", "^q9b", "^q9e")

t_test_results <- list()

for (pattern in competence) {
  matched_columns <- grep(pattern, colnames(allYrsFinal), value = TRUE)
  
  for (column in matched_columns) {
    t_test_result <- t.test(allYrsFinal[[column]] ~ allYrsFinal$gender)
    
    t_test_results[[column]] <- t_test_result
  }
}
t_test_results

```

## Per Question Visuals
```{r}
competenceFull <- c("q6e.pre.a", "q6e.post.a", "q7a.pre.a", "q7a.post.a", "q8f.pre.a", "q8f.post.a", "q9a.pre.a", "q9a.post.a", "q9b.pre.a", "q9b.post.a", "q9e.pre.a", "q9e.post.a")

questions_long <- allYrsFinal |>
  select(all_of(competenceFull), gender) |>
  pivot_longer(cols=all_of(competenceFull),
               names_to = "question",
               values_to = "score")
```
Pre Questions By Gender Boxplots
```{r}
questions_long|>
  group_by(gender)|>
  filter(str_detect(question, "pre")) |>
  ggplot(aes(x=question, y = score, color = gender)) + geom_boxplot() 
```

Post Questions by Gender Boxplots
```{r}
questions_long|>
  group_by(gender)|>
  filter(str_detect(question, "post")) |>
  ggplot(aes(x=question, y = score, color = gender)) + geom_boxplot() 
```

Pre vs. Post Faceted by Gender
```{r}
questions_long <- questions_long |>
  mutate(preOrPost = case_when(str_detect(question, "pre") ~ "pre",
                               str_detect(question, "post") ~ "post"),
         question_num = str_extract(question, "q\\d+[a-z]"))

questions_long|>
  group_by(preOrPost) |>
  ggplot(aes(x=question_num, y= score, color = preOrPost)) + 
  geom_boxplot() +
  facet_grid(.~gender) +
  scale_x_discrete(labels = function(x) str_extract(x, "\\d+[a-z]"))


```
Change Per Question by Gender
```{r}
questions_long <- questions_long |>
  mutate(row_id = row_number())

questions_wide <- questions_long |>
  pivot_wider(names_from = preOrPost, values_from = score, id_cols = c(row_id, gender, question_num)) |>
  select(-row_id) |>
  group_by(gender, question_num) |>
  fill(pre, post, .direction = "downup") |>
  ungroup()

questions_change <- questions_wide |>
  mutate(change = post - pre) 

questions_change|>
  ggplot(aes(x=question_num, y=change, color = gender)) + geom_boxplot() 
```



## Where there are "other" or "don't want to answer" for the gender question

```{r}
unique(allYrsFinal$gender)

allYrsFinal|>
  filter(is.na(gender))
```

### Mean pre cog comp by gender
```{r}
# note: NAs have the lowest pre and post cognitive competence, and have the greatest decrease in cognitive competence
allYrsFinal|>
  group_by(gender)|>
  summarize(mean(cognitive.competence.pre))
```

### Mean post cog comp by gender
```{r}
allYrsFinal|>
  group_by(gender)|>
  summarize(mean(cognitive.competence.post))
```

### Mean change cog comp by gender
```{r}
allYrsFinal|>
  group_by(gender)|>
  summarize(mean(cognitive.competence.change))
```

### How much do individual questions differ in terms of mean female - mean male to other ones within cog comp?
```{r}
competence <- c("^q6e", "^q7a", "^q8f", "^q9a", "^q9b", "^q9e")
competenceFull <- c("q6e.pre.a", "q6e.post.a", "q7a.pre.a", "q7a.post.a", "q8f.pre.a", "q8f.post.a", "q9a.pre.a", "q9a.post.a", "q9b.pre.a", "q9b.post.a", "q9e.pre.a", "q9e.post.a")
#competenceFull
qDiffsGender <- list()
for (q in competenceFull) {
  q <- sym(q)
  question <- allYrsFinal|>
    group_by(gender)|>
    filter(!is.na(!!q))|>
    summarize(mean(!!q))
  qDiffsGender[[q]] <- question[[1,2]] - question[[2,2]]
}

#questionDiffs

qDiffsGender_df <- enframe(unlist(qDiffsGender), name = "question", value = "diff")
qDiffsGender_df
```

## "Max Possible Score" on cogcomp.pre
overall max : 7
```{r}
max(allYrsFinal$cognitive.competence.pre)
```

## Is change in attitude is a better predictor of achievable than the actual attitude? 

Change in attitude
* change in attitudes is a significant predictor of achievable gain
* adj R^2 = 0.01189
* rmse = 0.2702
```{r}
change_ach.fit <- lm(ach.gain.24 ~ cognitive.competence.change, data = allYrsFinal)

summary(change_ach.fit)
```

Actual attitude (pre)
* pre-attitudes in cognitive competence is a statistically significant predictor of achievable gain
* adj R^2 = 0.00676
* rmse = 0.2709
```{r}
pre_ach.fit <- lm(ach.gain.24 ~ cognitive.competence.pre, data = allYrsFinal)

summary(pre_ach.fit)
```

Actual attitude (post)
* post attitudes in cognitive competence are a significant predictor of achievable gain
* adj R^2 = 0.03201
* rmse = 0.2674
```{r}
post_ach.fit <- lm(ach.gain.24 ~ cognitive.competence.post, data = allYrsFinal)

summary(post_ach.fit)
```

Actual attitude (pre AND post)
* model with pre and post attitudes in cognitive competence is a significant predictor of achievable gain
* adj R^2 = 0.03194
* rmse = 0.2675
```{r}
pre_post_ach.fit <- lm(ach.gain.24 ~ cognitive.competence.post + cognitive.competence.pre, data = allYrsFinal)

summary(pre_post_ach.fit)
```

### Comparing the models
Post-attitudes have the lowest AIC and BIC (best predictor)
```{r}
AIC(change_ach.fit, pre_ach.fit, post_ach.fit, pre_post_ach.fit)
```

```{r}
BIC(change_ach.fit, pre_ach.fit, post_ach.fit, pre_post_ach.fit)
```

## Interactions
With post-attitude as the response
* cognitive.competence.pre:cognitive.competence.change not significant
```{r}
# lm(y~x1+x2+x1:x2, data=my.data) 
post_int.fit <- lm(cognitive.competence.post ~ cognitive.competence.pre+cognitive.competence.change + cognitive.competence.pre:cognitive.competence.change, data = allYrsFinal)

summary(post_int.fit)
```

With achievable gain as the response
* cognitive.competence.pre:cognitive.competence.change not significant
```{r}
# lm(y~x1+x2+x1:x2, data=my.data) 
ach_int1.fit <- lm(ach.gain.24 ~ cognitive.competence.pre+cognitive.competence.change + cognitive.competence.pre:cognitive.competence.change, data = allYrsFinal)

summary(ach_int1.fit)
```