---
title: "Optional Subsetting"
author: Ajay Patel and Sean Leader
output: html_notebook
---

### Reading in the Data ###
```{r}
library(plyr)
library(tidyverse)
#library(here)

#### Reading in All Years Data - 14792 obs ####
#alldata <- read.csv(here::here("Data", "All Years Final Public.csv"), stringsAsFactors = TRUE)
alldata <- read.csv("All Years Final Public.csv", stringsAsFactors = TRUE)

data <- alldata
nrow(data)
```
### Keep students who only took Pre and Post ###
```{r}
### Subset of data where person consented on both pre and post ####
## 14792 to 11800: Note that due to previous subsetting, only year 3 has students who answered NA on Opt.out_pre/post ##

data <- data %>%
  filter(is.na(opt.out.pre) == F & is.na(opt.out.post) == F)
```


### Remove students with low response rate pre and post ###
```{r}
### c.rr.pre = proportion of non NA values for concept questions (Q17 on) in each row
data$post.perc.24
data$post.perc.33
data$pre.perc.24
data$pre.perc.33
c.rr.pre
c.rr.post


### 648 obs with low pre RR - alldata ###
### 34 obs with low pre RR - data (keeping students who only took pre and post) ###
data <- data[data$c.rr.pre >= .8,]

### 2449 obs with low post RR - alldata  ###
### 35 obs with low post RR - data (keeping students who only took pre and post)###
data <- data[data$c.rr.post >= .8,]

### If you do both of the above steps, removes 77 obs total
```


### Removes extreme achievable gain ###
```{r}
### c.post.cc

### ach_gain_33 = (post-pre)/(1-pre)
### ach_gain_24 uses the question sets

### alldata ###
### 674 obs with low ach_gain_24 (-1.1)
### 89 obs wtih low ach_gain_24 (-2)
### 606 with low ach_gain_33 (-1.1)
### 92 obs with low ach_gain_33 (-2)

### After subsetting students who only took pre and post ###
### 13 obs with low ach_gain_24 (-1.1) 
### 2 obs with low ach_gain_24 (-2) 
### 7 obs with low ach_gain_33 (-1.1) 
### 2 obs with low ach_gain_33 (-2)

data <- data[data$ach.gain.24 > -1.1,]
```

### Remove ISCAM and own ###
```{r}
# #### Remove ISCAM students ####
# ### Lose up to 113 obs by taking out ISCAM - alldata ###
# ### Lose up to 111 obs by taking out ISCAM - data (keeping students who only took pre and post) ###
# data <- data[data$textbook.classification != "ISCAM",]
# data$textbook.classification <- droplevels(data$textbook.classification)
# 
# # ### Re-orders if you remove only ISCAM  ###
# # data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2", "Other"))
# 
# 
# #### Remove instructors using "own" materials ####
# ### Lose up to 344 obs by taking out own - alldata ###
# ### Lose up to 268 obs by taking out own - data (keeping students who only took pre and post) ###
# data <- data[data$textbook.classification != "Other",]
# data$textbook.classification <- droplevels(data$textbook.classification)
 
# # ### Re-orders if you remove only own  ###
# # data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2", "ISCAM"))
# 
# 
# ### Re-orders if you remove both ISCAM and own ###
# data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2"))
```


### Remove calculus math.prereq ###
```{r}
### Lose up to 513 obs by taking out calculus - alldata ###
### Lose up to 425 obs by taking out calculus - data (keeping only students who only took pre and post) ###
### Note, that removing Calculus prereq also removes all the ISCAM from textbook.classification ###
data <- data %>%
  filter(math.prereq != "Calculus" | is.na(math.prereq) == T)

data$math.prereq <- droplevels(data$math.prereq)
data$textbook.classification <- droplevels(data$textbook.classification)
```


#### RUN LINE TO TAKE OUT HS STUDENTS (Only based on Carnegie) ####
```{r}
### 1466 HS students - alldata  ###
### 793- data (keeping students who only took pre and post) ###
data <- data %>%
  filter(carnegie.classification != "High School")

### NOTE, when removing HS student both Carnegie and student.type need to be refactored ###
### There are no more "High School Students" in student.type ###
data$carnegie.classification <- droplevels(data$carnegie.classification)
data$student.type <- droplevels(data$student.type)

### Re-orders the factors ###
data$carnegie.classification <- factor(data$carnegie.classification, levels = c("Community College", "Baccalaureate College", "Master's", "Doctoral Universities"))
```


#### Take out class size (instructor reported at end of course) less than 9 ####
```{r}
### 57 obs have small sections (leaving NAs in) - alldata ###
### 324 obs have small sections and NAs - alldata ###

### 35 obs have small sections (leaving NAs in) - alldata ###
### 302 obs have small sections and NAs - alldata ###

#suggest leaving in NAs

#data <- data[(data$class.size.end > 9 | is.na(data$class.size.end) == T), ]
```

### Checks the number of NAs in each column by year ###
```{r}
data %>%
  group_by(year) %>%
  summarise_all(funs(sum(is.na(.))))
```

### Writing out data ###
```{r}
write.csv(data, "Post Optional Subsetting.csv")
```




