---
title: "Optional Subsetting"
author: Ajay Patel
output: html_notebook
---



### Reading in the Data ###
```{r}
library(plyr)
library(tidyverse)

# Added 5/2 by Sean Leader:
library(here)
#

load(url("http://www.rossmanchance.com/iscam3/ISCAM.RData"))

#### Reading in All Years Data - 14699 obs ####

# Original read-in message, replaced with code below
#alldata <- read.csv("/Users/ajaypatel21/Desktop/Frost/All Years Final v3 Public.csv", stringsAsFactors #= T)

# Added 5/2 by Sean Leader:
alldata <- read.csv(here::here("Data", "All Years Final v3 Public.csv"), stringsAsFactors = TRUE)

data <- alldata
```


```{r}
# Moved 5/2 by Sean Leader to separate code chunk for clarity

### Textbook Classification Table ####
### Shows how which textbooks got classified ###
# textbook_table <- data %>%
#   select(textbook.used, textbook.classification, year) %>%
#   group_by(textbook.used, textbook.classification, year) %>%
#   tally()
 
# write.csv(textbook_table, "Data/Textbook Classification Table.csv", row.names = F)
```




### Keep students who only took Pre and Post ###
```{r}
### Subset of data where person consented on both pre and post ####
## 14792 to 11800: Note that due to previous versions of data that have already been subsetted, only year 3 has ##
## students who answered NA on Opt.out_pre/post ##

data <- data %>%
  filter(is.na(opt.out.pre) == F & is.na(opt.out.post) == F)
```


### Remove low response rate pre and post ###
```{r}
### Lose up to 648 obs removing low pre RR - alldata ###
### Lose up to 34 obs removing low pre RR - data (keeping students who only took pre and post) ###
data <- data[data$c.rr.pre >= .8,]

### Lose up to 2449 obs by removing low post RR - alldata  ###
### Lose up to 35 obs by removing low post RR - data (keeping students who only took pre and post)###
data <- data[data$c.rr.post >= .8,]

### If you do both of the above steps, you should have lost 77 obs total
```


### Removes extreme achievable gain ###
```{r}
### Lose up to 674 obs by removing low ach_gain_24 (-1.1) - alldata ###
### Lose up to 89 obs by removing low ach_gain_24 (-2) - alldata ###
### Lose up to 606 obs by removing low ach_gain_33 (-1.1) - alldata ###
### Lose up to 92 obs by removing low ach_gain_33 (-2) - alldata ###

### Lose up to 13 obs by removing low ach_gain_24 (-1.1) - data (keeping students who only took pre and post) ###
### Lose up to 2 obs by removing low ach_gain_24 (-2) - data (keeping students who only took pre and post) ###
### Lose up to 7 obs by removing low ach_gain_33 (-1.1) - data (keeping students who only took pre and post) ###
### Lose up to 2 obs by removing low ach_gain_33 (-2) - data (keeping students who only took pre and post) ###

data <- data[data$ach.gain.24 > -1.1,]
```


### Remove ISCAM and own ###
```{r}
# #### Remove ISCAM students ####
# ### Lose up to 113 obs by taking out ISCAM - alldata ###
# ### Lose up to 111 obs by taking out ISCAM - data (keeping students who only took pre and post) ###
# data <- data[data$textbook.classification != "ISCAM",]
# data$textbook.classification <- droplevels(data$textbook.classification)
# 
# # ### Re-orders if you remove only ISCAM  ###
# # data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2", "Other"))
# 
# 
# #### REMOVE OWN PEOPLE ####
# ### Lose up to 344 obs by taking out own - alldata ###
# ### Lose up to 268 obs by taking out own - data (keeping students who only took pre and post) ###
# data <- data[data$textbook.classification != "Other",]
# data$textbook.classification <- droplevels(data$textbook.classification)
# 
# # ### Re-orders if you remove only own  ###
# # data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2", "ISCAM"))
# 
# 
# ### Re-orders if you remove both ISCAM and own ###
# data$textbook.classification <- factor(data$textbook.classification, levels = c("ISI", "ISI1st", "OtherSBI", "NotSBI", "NotSBI2"))
```


### Remove calculus math.prereq ###
```{r}
### Lose up to 513 obs by taking out calculus - alldata ###
### Lose up to 425 obs by taking out calculus - data (keeping only students who only took pre and post) ###
### Note, that removing Calculus prereq also removes all the ISCAM from textbook.classification ###
data <- data %>%
  filter(math.prereq != "Calculus" | is.na(math.prereq) == T)

data$math.prereq <- droplevels(data$math.prereq)
data$textbook.classification <- droplevels(data$textbook.classification)
```


#### RUN LINE TO TAKE OUT HS STUDENTS (Only based on Carnegie) ####
```{r}
### Lose up to 1466 obs by taking out HS students - alldata  ###
### Lose up to 793 obs by taking out HS students - data (keeping students who only took pre and post) ###
data <- data %>%
  filter(carnegie.classification != "High School")

### NOTE, when removing HS student both Carnegie and student.type need to be refactored ###
### There are no more "High School Students" in student.type ###
data$carnegie.classification <- droplevels(data$carnegie.classification)
data$student.type <- droplevels(data$student.type)

### Re-orders the factors ###
data$carnegie.classification <- factor(data$carnegie.classification, levels = c("Community College", "Baccalaureate College", "Master's", "Doctoral Universities"))
```


#### Take out class size less than 9 ####
```{r}
### Lose up to 57 obs by taking out reported small sections (leaving NAs in) - alldata ###
### Lose up to 324 obs by taking out small sections and NAs - alldata ###

### Lose up to 35 obs by taking out reported small sections (leaving NAs in) - alldata ###
### Lose up to 302 obs by taking out small sections and NAs - alldata ###

# Uncommented on 5/2 by Sean Leader 
data <- data[(data$class.size.end > 9 | is.na(data$class.size.end) == T), ]
```


### Remove low section c RR ###
```{r}
### Lose up to 2007 obs by removing sections with low rr ###

# No code currently available to remove low section response rate
```


### Checks the number of NAs in each column by year ###
```{r}
data %>%
  group_by(year) %>%
  summarise_all(funs(sum(is.na(.))))
```


### Amelia Package ###
```{r}
### NOTE, Amelia imputations do not work with factors and characters - have to change to numeric ###
### NOTE, Amelia should be run with data and not alldata due to high volume of missing values and students who don't take pre or post test ###

library(Amelia)

### Imputes student mean attitudes, demographics, GPA, math_satact_flag, SATACTzscore, overall attitude ###
### NOTE, I am not using firstgen, race_origin, and is.white since these variables would have been imputed but they were not actually asked in year 1, hence the values should be missing ###

data_amelia <- data

### Changes factor variables to character ###
data_amelia$is.major.course <- as.character(data_amelia$is.major.course)
data_amelia$field.major <- as.character(data_amelia$field.major)
data_amelia$grade.expectation.pre <- as.character(data_amelia$grade.expectation.pre)
data_amelia$grade.expectation.post <- as.character(data_amelia$grade.expectation.post)
data_amelia$gender <- as.character(data_amelia$gender)
data_amelia$has.taken.stat.course <- as.character(data_amelia$has.taken.stat.course)
data_amelia$prev.stat.course <- as.character(data_amelia$prev.stat.course)
data_amelia$status <- as.character(data_amelia$status)
data_amelia$math.satact.flag <- as.character(data_amelia$math.satact.flag)

### Changes the levels of variables to numeric ###
data_amelia$is.major.course[data_amelia$is.major.course == "major course"] <- 1
data_amelia$is.major.course[data_amelia$is.major.course == "not major course"] <- 0
data_amelia$field.major[data_amelia$field.major == "Arts and humanities"] <- 1
data_amelia$field.major[data_amelia$field.major == "Natural and applied sciences"] <- 2
data_amelia$field.major[data_amelia$field.major == "Other"] <- 3
data_amelia$field.major[data_amelia$field.major == "Social sciences"] <- 4

data_amelia$grade.expectation.pre <- mapvalues(data_amelia$grade.expectation.pre,
                                                  from = c("A+", "A", "A-", "B+", "B", "B-","C+", "C", "C-",
                                                           "D+", "D", "D-", "F"),
                                                  to = c(1:13))


data_amelia$grade.expectation.post <- mapvalues(data_amelia$grade.expectation.post,
                                               from = c("A+", "A", "A-", "B+", "B", "B-",
                                                      "C+", "C", "C-", "D+", "D", "D-", "F"),
                                               to = c(1:13))

data_amelia$gender[data_amelia$gender == "Female"] <- 1
data_amelia$gender[data_amelia$gender == "Male"] <- 0
data_amelia$has.taken.stat.course[data_amelia$has.taken.stat.course == "No"] <- 0
data_amelia$has.taken.stat.course[data_amelia$has.taken.stat.course == "Yes"] <- 1
data_amelia$prev.stat.course[data_amelia$prev.stat.course == "No"] <- 0
data_amelia$prev.stat.course[data_amelia$prev.stat.course == "Yes (Not AP)"] <- 1
data_amelia$prev.stat.course[data_amelia$prev.stat.course == "Yes (AP)"] <- 2

data_amelia$status <- mapvalues(data_amelia$status,
                               from = c("Other", "High School Student", "Freshman in college", "Sophomore in college",
                                      "Junior in college", "Senior in college", "Fifth or more year in college",
                                      "Non-traditional/part-time student in college", "Graduate Student"),
                               to = c(0:8))

data_amelia$math.satact.flag[data_amelia$math.satact.flag == "SAT"] <- 1
data_amelia$math.satact.flag[data_amelia$math.satact.flag == "ACT"] <- 2

### Ensures changed variables are numeric and not character/factor ###
data_amelia$is.major.course <- as.numeric(data_amelia$is.major.course)
data_amelia$field.major <- as.numeric(data_amelia$field.major)
data_amelia$grade.expectation.pre <- as.numeric(data_amelia$grade.expectation.pre)
data_amelia$grade.expectation.post <- as.numeric(data_amelia$grade.expectation.post)
data_amelia$gender <- as.numeric(data_amelia$gender)
data_amelia$has.taken.stat.course <- as.numeric(data_amelia$has.taken.stat.course)
data_amelia$prev.stat.course <- as.numeric(data_amelia$prev.stat.course)
data_amelia$status <- as.numeric(data_amelia$status)
data_amelia$math.satact.flag <- as.numeric(data_amelia$math.satact.flag)


### math.perf.pre through field.major for amelia imputing ###
ord_vars <- names(data_amelia[c(225:238)])

### Creates a bound so that min gpa = 1.0 and max gpa = 4.0 ###
gpa_bound <- matrix(c(239, 1.0, 4.0), nrow = 1, ncol = 3)

### Runs the amelia imputation - should not have any missing values in this subset of data ###
### first argument = dataset, m = number of imputations/outputted data sets, bounds = variables that have any bounds ###
### noms = binary variables, ords = factor variables as numeric, idvars = variables that will be outputted but will not ###
### be imputed ###

### X to q9.f.post.a, firstgen to year, days.meet to q4041.post.c, c.rr.pre to a.rr.post, pre.perc.33 to d ###
myidvars <- names(data_amelia[c(1:224, 247:266, 268:318, 331:334, 337:387)])

am.out <- amelia(data_amelia, m=1,
                 idvars = myidvars,
                 bounds = gpa_bound,
                 noms = c("gender", "has.taken.stat.course"),
                 ords = c(ord_vars, "grade.expectation.pre", "grade.expectation.post",
                          "prev.stat.course", "status", "math.satact.flag"))

### Saves as a new CSV file ###

# Changed 5/2 by Sean Leader to go to a Data folder in specific.
write.amelia(am.out, file.stem = "Data/outdata")
```


### Re-reverse Amelia code and write out ###
```{r}

# Changed 5/2 by Sean Leader to use the here package to load the previous dataset
data_imputed <- read.csv(here::here("Data", "outdata1.csv"))

data_imputed$gpa <- round(data_imputed$gpa, 2)

### Changes the levels of variables from numeric to factor ###
data_imputed$is.major.course[data_imputed$is.major.course == 1] <- "major course"
data_imputed$is.major.course[data_imputed$is.major.course == 0] <- "not major course"
data_imputed$field.major[data_imputed$field.major == 1] <- "Arts and humanities"
data_imputed$field.major[data_imputed$field.major == 2] <- "Natural and applied sciences"
data_imputed$field.major[data_imputed$field.major == 3] <- "Other"
data_imputed$field.major[data_imputed$field.major == 4] <- "Social sciences"

data_imputed$grade.expectation.pre <- mapvalues(data_imputed$grade.expectation.pre,
                                                from = c(1:13),
                                                to = c("A+", "A", "A-", "B+", "B", "B-","C+", "C", "C-",
                                                         "D+", "D", "D-", "F"))


data_imputed$grade.expectation.post <- mapvalues(data_imputed$grade.expectation.post,
                                                 from = c(1:13),
                                                 to = c("A+", "A", "A-", "B+", "B", "B-",
                                                      "C+", "C", "C-", "D+", "D", "D-", "F"))

data_imputed$gender[data_imputed$gender == 1] <- "Female"
data_imputed$gender[data_imputed$gender == 0] <- "Male"
data_imputed$has.taken.stat.course[data_imputed$has.taken.stat.course == 0] <- "No"
data_imputed$has.taken.stat.course[data_imputed$has.taken.stat.course == 1] <- "Yes"
data_imputed$prev.stat.course[data_imputed$prev.stat.course == 0] <- "No"
data_imputed$prev.stat.course[data_imputed$prev.stat.course == 1] <- "Yes (Not AP)"
data_imputed$prev.stat.course[data_imputed$prev.stat.course == 2] <- "Yes (AP)"

data_imputed$status <- mapvalues(data_imputed$status,
                                 from = c(0:8),
                                 to = c("Other", "High School Student", "Freshman in college", "Sophomore in college",
                                      "Junior in college", "Senior in college", "Fifth or more year in college",
                                      "Non-traditional/part-time student in college", "Graduate Student"))

data_imputed$is.white[data_imputed$is.white == 1] <- T
data_imputed$is.white[data_imputed$is.white == 0] <- F

data_imputed$math.satact.flag[data_imputed$math.satact.flag == 1] <- "SAT"
data_imputed$math.satact.flag[data_imputed$math.satact.flag == 2] <- "ACT"


### Ensures changed variables are numeric and not character/factor ###
data_imputed$is.major.course <- as.factor(data_imputed$is.major.course)
data_imputed$field.major <- as.factor(data_imputed$field.major)
data_imputed$grade.expectation.pre <- as.factor(data_imputed$grade.expectation.pre)
data_imputed$grade.expectation.post <- as.factor(data_imputed$grade.expectation.post)
data_imputed$gender <- as.factor(data_imputed$gender)
data_imputed$has.taken.stat.course <- as.factor(data_imputed$has.taken.stat.course)
data_imputed$prev.stat.course <- as.factor(data_imputed$prev.stat.course)
data_imputed$status <- as.factor(data_imputed$status)
data_imputed$math.satact.flag <- as.factor(data_imputed$math.satact.flag)

### Since mean attitudes were imputed, need to recalculate attitude change variables ###
data_imputed$affect.change <- data_imputed$affect.post - data_imputed$affect.pre
data_imputed$cognitive.competence.change <- data_imputed$cognitive.competence.post - data_imputed$cognitive.competence.pre
data_imputed$difficulty.change <- data_imputed$difficulty.post - data_imputed$difficulty.pre
data_imputed$effort.change <- data_imputed$effort.post - data_imputed$effort.pre
data_imputed$interest.change <- data_imputed$interest.post - data_imputed$interest.pre
data_imputed$value.change <- data_imputed$value.post - data_imputed$value.pre

### Need to recalculate section level variables because of imputation ### 
data_imputed$answered.pre <- 1
data_imputed$answered.pre[data_imputed$pre.perc.24 == 0] <- 0

data_imputed$answered.post <- 1
data_imputed$answered.post[data_imputed$post.perc.24 == 0] <- 0

data_imputed$answered.both <- 1
data_imputed$answered.both[data_imputed$pre.perc.24 == 0 | data_imputed$post.perc.24 == 0] <- 0

data_imputed <- data_imputed %>%
  group_by(instructor.section, year) %>%
  mutate(section.num.responded = n(), 
         section.pre.responded = sum(answered.pre, na.rm = T),
         section.post.responded = sum(answered.post, na.rm = T),
         section.both.responded = sum(answered.both, na.rm = T),
         # section.c.rr1 = section.both.responded/section.pre.responded, 
         # section.c.rr2 = section.post.responded/section.pre.responded,
         # section.pre.rr = section.pre.responded/class.size.start,
         # section.post.rr = section.post.responded/class.size.end,
         # section.pre.rr.new = section.pre.responded/section.num.responded,
         # section.post.rr.new = section.post.responded/section.num.responded,
         section.c.rr = section.both.responded/class.size.start, 
         section.ach.gain.24 = mean(ach.gain.24, na.rm = T),
         section.pre.perc.24 = mean(pre.perc.24, na.rm = T),
         section.post.perc.24 = mean(post.perc.24, na.rm = T),         
         section.gpa = mean(gpa, na.rm = T),
         section.satact.zscore = mean(satact.zscore, na.rm = T),
         section.affect.pre = mean(affect.pre, na.rm = T),
         section.affect.post = mean(affect.post, na.rm = T),
         section.cognitive.competence.pre = mean(cognitive.competence.pre, na.rm = T),
         section.cognitive.competence.post = mean(cognitive.competence.post, na.rm = T),
         section.difficulty.pre = mean(difficulty.pre, na.rm = T),
         section.difficulty.post = mean(difficulty.post, na.rm = T),
         section.effort.pre = mean(effort.pre, na.rm = T),
         section.effort.post = mean(effort.post, na.rm = T),
         section.interest.pre = mean(interest.pre, na.rm = T),
         section.interest.post = mean(interest.post, na.rm = T),
         section.value.pre = mean(value.pre, na.rm = T),
         section.value.post = mean(value.post, na.rm = T),
         section.overall.attitude.pre = mean(overall.attitude.pre, na.rm = T),
         section.overall.attitude.post = mean(overall.attitude.post, na.rm = T),
         var.pre = var(pre.perc.24, na.rm = T),
         var.post = var(post.perc.24, na.rm = T)
         ) %>%
  ungroup() %>%
  mutate(d = (section.post.perc.24 - section.pre.perc.24) / sqrt((var.pre+var.post)/2))


### Changing NA's to Unknown ###
data_imputed$class.session.length[is.na(data_imputed$class.session.length) == T] <- "Unknown"
data_imputed$class.size.start.ind[is.na(data_imputed$class.size.start.ind) == T] <- "Unknown"
data_imputed$class.size.end.ind[is.na(data_imputed$class.size.end.ind) == T] <- "Unknown"
data_imputed$class.meet.weeks.ind[is.na(data_imputed$class.meet.weeks.ind) == T] <- "Unknown"
data_imputed$time.meet[is.na(data_imputed$time.meet) == T] <- "Unknown"
data_imputed$incentive.pre[is.na(data_imputed$incentive.pre) == T] <- "Unknown"
data_imputed$incentive.post[is.na(data_imputed$incentive.post) == T] <- "Unknown"
data_imputed$math.prereq[is.na(data_imputed$math.prereq) == T] <- "Unknown"
data_imputed$student.type[is.na(data_imputed$student.type) == T] <- "Unknown"
data_imputed$firstgen[is.na(data_imputed$firstgen) == T] <- "Unknown"
data_imputed$race.origin[is.na(data_imputed$race.origin) == T] <- "Unknown"
data_imputed$is.white[is.na(data_imputed$is.white) == T] <- "Unknown"
data_imputed$gaise.familiar[is.na(data_imputed$gaise.familiar) == T] <- "Unknown"
data_imputed$department.type[is.na(data_imputed$department.type) == T] <- "Unknown"
data_imputed$is.stats.department[is.na(data_imputed$is.stats.department) == T] <- "Unknown"
data_imputed$analyzing.data.experience[is.na(data_imputed$analyzing.data.experience) == T] <- "Unknown"
data_imputed$position.classification[is.na(data_imputed$position.classification) == T] <- "Unknown"
data_imputed$advanced.stats.degree.type[is.na(data_imputed$advanced.stats.degree.type) == T] <- "Unknown"
data_imputed$lecture.type[is.na(data_imputed$lecture.type) == T] <- "Unknown"
data_imputed$TA[is.na(data_imputed$TA) == T] <- "Unknown"
data_imputed$admin.pre.location[is.na(data_imputed$admin.pre.location) == T] <- "Unknown"
data_imputed$admin.post.location[is.na(data_imputed$admin.post.location) == T] <- "Unknown"
data_imputed$admin.pre.time[is.na(data_imputed$admin.pre.time) == T] <- "Unknown"
data_imputed$admin.post.time[is.na(data_imputed$admin.post.time) == T] <- "Unknown"
data_imputed$isi.workshop[is.na(data_imputed$isi.workshop) == T] <- "Unknown"
data_imputed$days.meet[is.na(data_imputed$days.meet) == T] <- "Unknown"
data_imputed$years.teaching.intro.stats.binned[is.na(data_imputed$years.teaching.intro.stats.binned) == T] <- "Unknown"
data_imputed$years.teaching.experience.binned[is.na(data_imputed$years.teaching.experience.binned) == T] <- "Unknown"
data_imputed$percent.lecture.binned[is.na(data_imputed$percent.lecture.binned) == T] <- "Unknown"

# Changed 5/2 by Sean Leader to go to a data folder
write.csv(data_imputed, "Data/Post Optional Subsetting Imputed.csv")
```


### Writing out non-Imputed Data ###
```{r}
# Changed 5/2 by Sean Leader to go to a data folder
write.csv(data, "Data/Post Optional Subsetting.csv")
```




